Demonstration of a Standardized Methodology for Building and Using Individual Movement Models using Chinook Salmon Pop-up Archival Tagging Data from the Gulf of Alaska

### 2024-09-26 proposal/Proposal.tex
edits

{\Huge \bf \@title\par}
\vspace{2.5cm}

@@

\vspace{2.5cm}
{\Large Dr. Andrew Seitz}\hspace{2cm}{\Large Dr. Curry Cunningham}\\[2cm]{\Large Michael Courtney, M.S.}\\[2cm]
College of Fisheries and Ocean Sciences\\
University of Alaska Fairbanks

@@

\title{Master's Research Proposal (DRAFT)}

@@

\section{Summary}

@@

The ultimate goal of this research is to provide a new pathway by which collections of high resolution spatio-temporal movement data that are representative of a species can be used to make a direct impact on fisheries management. It is hoped that this will provide an incentive for creating/enlarging such datasets.

\begin{objective}
To provide tooling that allows for using machine learning methods to fit models of the form

@@

that maximize the following objective:

@@

These models will be known as odds models as they predict the "odds for" each outcome $v_k$ given the information contained in $\eta_k$.

@@

\begin{enumerate}
\item \textbf{Understanding Connectivity:} by looking at the vector flows of fish throughout the space we will be able to identify which areas are highly connected and which are more or less separated by environment, cohort, or other factors of fish behavior.
\item \textbf{Understanding Vulnerability:} by especially looking at changes in depth occupancy and preference for habitat we will attempt to quantify the degree of vulnerability to different kinds of activities as a function of space and time. Specific vulnerabilities of interest might include: vulnerability to specific fishing fleets, vulnerability to naval training activities, vulnerability to marine mammals. 
\item \textbf{Cumulative Mortality:} by being able to follow cohorts through space and time we should be able to show how different fishing pressure scenarios might affect the cumulative fishing mortality. Note that the data for specific fleets may be unavailable but that our interest here is to demonstrate that given data we can make these kinds of predictions.  
\item \textbf{Protected Areas/Times:} Of those fishing pressure scenarios, one of especial interest is modeling Marine Protected Areas (MPAs) or other kinds of closures in space and time. Given the motility of Chinook, how effective would marine protected areas be? 
\item \textbf{Uncertainty:} we will show that such a simulation can be run at scale, bootstrapping can be applied, and therefore uncertainty in simulation outcomes quantified. 
\end{enumerate}

>>>>

{\Large \bf \@title\par}
\vspace{1cm}

{\Large \bf (Proposal) \par}
\vspace{2.5cm}

@@

\vspace{1cm}
{\Large Dr. Andrew Seitz}\hspace{2cm}{\Large Dr. Curry Cunningham}\\[2cm]{\Large Michael Courtney, M.S.}\\[1cm]
College of Fisheries and Ocean Sciences\\
University of Alaska Fairbanks

@@

\title{Demonstration of a Standardized Methodology for Building and Using Individual Movement Models using Chinook Salmon Pop-up Archival Tagging Data from the Gulf of Alaska}

@@

\section{Purpose}

@@

\begin{objective}
Movement modeling is fundamentally about modeling the selection of movement outcomes. So, given a set of choices $v_k$ and a set of features $n_k$ describing those choices, we want a standardized method to fit, evaluate, and use models of the form:

@@

where $\psi_k$ is the "odds for" choice $v_k$. Such a model should maximize the following objective:

@@

@@

\begin{enumerate}
\item \textbf{Understanding Connectivity:} by looking at the vector flows of fish throughout the space we will be able to identify which areas are highly connected and which are more or less separated by environment, cohort, or other factors of fish behavior.
\item \textbf{Understanding Vulnerability:} by especially looking at changes in depth occupancy and preference for habitat we will attempt to quantify the degree of vulnerability to different kinds of activities as a function of space and time. Specific vulnerabilities of interest might include: vulnerability to specific fishing fleets, vulnerability to naval training activities, vulnerability to marine mammals. 
\item \textbf{Cumulative Mortality:} by being able to follow cohorts through space and time we should be able to show how different fishing pressure scenarios might affect the cumulative fishing mortality. Note that the data for specific fleets may be unavailable but that our interest here is to demonstrate that given data we can make these kinds of predictions.  
\item \textbf{Protected Areas/Times:} Of those fishing pressure scenarios, one of especial interest is modeling Marine Protected Areas (MPAs) or other kinds of closures in space and time. Given the motility of Chinook, how must effective protected areas/times be designed? 
\item \textbf{Uncertainty:} we will show that such a simulation can be run at scale, bootstrapping can be applied, and therefore uncertainty in simulation outcomes quantified. 
\end{enumerate}
